---
title: "PFAS_analysis_KN"
author: "Karly Nocera and Tay Holiday"
date: "4/16/2021"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
# Set your working directory

getwd()

# Load your packages

library(tidyverse)
library(lubridate)

# Set your ggplot theme

mytheme <- theme_classic(base_size = 14) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "top")

theme_set(mytheme)

# Load your datasets
PFAS_WWTP_clean <- read.csv("./Data/Processed/PFAS_WWTP_clean.csv", stringsAsFactors = TRUE)

PFAS_WWTP_wide_clean <- read.csv("./Data/Processed/PFAS_WWTP_wide_clean.csv", stringsAsFactors = TRUE)

PFAS_Source_clean <- read.csv("./Data/Processed/PFAS_Source_clean.csv", stringsAsFactors = TRUE)

PFAS_Source_wide_clean <- read.csv("./Data/Processed/PFAS_Source_wide_clean.csv", stringsAsFactors = TRUE)

```


# Analysis

Insert visualizations and text describing your main analyses. Format your R chunks so that graphs are displayed but code and other output is not displayed. Instead, describe the results of any statistical tests in the main text (e.g., "Variable x was significantly different among y groups (ANOVA; df = 300, F = 5.55, p < 0.0001)"). Each paragraph, accompanied by one or more visualizations, should describe the major findings and how they relate to the question and hypotheses. Divide this section into subsections, one for each research question.

Each figure should be accompanied by a caption, and each figure should be referenced within the text




## Question 1: Source Data Analytes

### Question 1a: Long v Short

```{r}

# want to know what type of remediation is needed to deal with their contamination (GAC short chain; RO both)

ggplot(subset(PFAS_WWTP_clean, !is.na(ppt)), aes(x = Site, y = ppt))+
  geom_col(aes(fill = chain.length)) +
  labs(fill = "Chain Length") +
  xlab("WWTP Site") +
  theme(axis.text.x = element_text(angle = 90,  hjust = 1))


ggplot(subset(PFAS_Source_clean, !is.na(ppt)), aes(x = Site, y = ppt))+
  geom_col(aes(fill = chain.length)) +
  labs(fill = "Chain Length") +
  xlab("Source Site") +
  theme(axis.text.x = element_text(angle = 90,  hjust = 1))



PFAS_wide_clean <- 
    PFAS_wide_clean %>% 
    rowwise() %>% 
    mutate(short.ppt = sum(PFBA, PFBS, PFHpA, PFHxA, PFPeS, PFPeA, na.rm = TRUE)) %>%
    mutate(long.ppt = sum(PFDS, PFDA, PFDoA, PFHpS, PFHxS, PFNS, PFNA, PFOSA, N.EtFOSAA, N.MeFOSAA, PFTeA, PFTriA, PFUnA, HFPO.DA, PFOA, PFOS, na.rm = TRUE))
  
# stacked histogram ppt vs site

ggplot(PFAS_WWTP_wide_clean, aes(x = Site, y = short.ppt))+
  geom_col(aes(fill = supp), width = 0.7)


ggplot(PFAS_WWTP_clean) +
  geom_col(aes(x = Site, y = ppt)) +
  scale_y_continuous(limits = c(0, 100))


PFAS_wide_clean %>% 
   select(-total) %>% 
   gather(type, count, burglaries:robberies) %>% 
   ggplot(., aes(x=year, y=count, fill=forcats::fct_rev(type))) +
   geom_bar(stat="identity")


```

### Question 1b: Top 5 Analytes

```{r}

# plot analytes by site shows the only analytes worth knowing

ggplot(PFAS_WWTP_clean, aes(fill=Analyte, y=ppt, x=Site)) +     geom_bar(position="stack", stat="identity") +
    labs(fill = "Analyte") +
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  xlab("Source Sites")

```


### Question 1c: Are there rivers that are scary we aren't talking about? (high and type of analyte)

```{r}

# PROBLEM: wide dataset shows 8 sites with PFAS, long shows 11

# decide top sites to plot

topsites <-
  PFAS_wide_clean %>%
  filter(Type =="Source") %>%
  select(Type, Site, TotalPFAS)


topsites <- topsites[with(topsites,order(-TotalPFAS)),]

topsites <- topsites[1:8,]


topsites_wwtp <-
  PFAS_WWTP_wide_clean %>%
  select(Site, TotalPFAS)

topsites_wwtp <- topsites_wwtp[with(topsites_wwtp,order(-TotalPFAS)),]

topsites_wwtp <- topsites_wwtp[1:20,]



# facet wrap each site by x = analyte y = ppt

ggplot(topsites_wwtp, aes(x = Site, y = TotalPFAS)) + 
  geom_bar(stat="identity") +
  facet_wrap(vars(Site), nrow = 5) +
  ylab("Total PFAS (ppt)") +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) 


ggplot(PFAS_WWTP_wide_clean, aes(x=month, y = TotalPFAS, group=Site)) +
  geom_line(aes(color=Site)) +
  geom_point(aes(color=Site)) +
  scale_x_continuous(breaks=c(7, 8, 9)) +
  ylim(0,600) +
  ylab("Total PFAS (ppt)") +
  xlab("2019 Month") +
  theme(legend.position="right")

```

### Question 1d: [Tay] Are any spots with one analyte over 70? over 30?

```{r}
#Any locations with Analyte above 30? 70?
#To do this, a simple pipe should do to filter out which analytes/site are above the given value.
#EPA_Processed_2018$Result <- subset(EPA_Processed_2018$Result, !is.na()) #converts the result into a numeric value instead of a factor

source_analyte_above_30 <- 
  PFAS_Source_clean %>%
  filter(ppt > 30)

source_analyte_above_70 <- 
  PFAS_Source_clean %>%
  filter(ppt > 70)

unique(source_analyte_above_30$Site) #11 Sites above 30 
unique(source_analyte_above_70$Site) #5 sites above 70


#Plotting the sites with more than 30/70
source_analyte_30_plot <- 
  ggplot(source_analyte_above_30, aes(x = Sample.Date, y = ppt, color = Site)) +
  geom_point(alpha = 0.5, size = 3)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  theme(axis.text.x = element_text(size = 7))+
  geom_hline(yintercept = 30, lty = 2)+
  scale_y_continuous(breaks=c(30, 100, 200, 300, 400, 500, 600))+
  theme(legend.position = "right", 
        legend.text = element_text(size = 5), legend.title = element_text(size = 12))+
  xlab("Sample Date")
print(source_analyte_30_plot + ggtitle("Source sites with Analytes above 30 ppt"))

source_analyte_70_plot <- 
  ggplot(source_analyte_above_70, aes(x = Sample.Date, y = ppt, color = Site)) +
  geom_point(alpha = 0.5, size = 3)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  theme(axis.text.x = element_text(size = 7))+
  geom_hline(yintercept = 70, lty = 2)+
  scale_y_continuous(breaks=c(70, 100, 200, 300, 400, 500, 600))+
  theme(legend.position = "right", 
        legend.text = element_text(size = 5), legend.title = element_text(size = 12))+
  xlab("Sample Date")
print(source_analyte_70_plot + ggtitle("Source sites with Analytes above 70 ppt"))

```


### Question 1e: [Tay] Total PFAS over 130

```{r}

#Total PFAS over 130

Total_PFAS_WWTP_Over130 <- 
  PFAS_WWTP_wide_clean %>%
  filter(TotalPFAS > 130)


# sum of pfas for every source site at that date
# filter for those over 130
# unique for how many sites had total pfas over 130

#Source by Site and Date with sum PFAS above 130
# Total_PFAS_Source_Over130 <- 
#   PFAS_long_clean[!is.na(PFAS_long_clean$ppt),] %>%
#   filter (Type == "Source") %>%
#   filter(Analyte == "PFBA"| Analyte == "PFBS" | Analyte == "PFDS" | Analyte == "PFDA" | Analyte == "PFDoA" | Analyte == "PFHpS" | Analyte == "PFHpA" | Analyte == "PFHxS" | Analyte == "PFHxA" | Analyte == "PFNS" | Analyte == "PFNA" | Analyte == "PFOSA" | Analyte == "N.EtFOSAA" | Analyte == "N.MeFOSAA" | Analyte == "PFPeS" | Analyte == "PFPeA" | Analyte == "PFTeA" | Analyte == "PFTriA" | Analyte == "PFUnA" | Analyte == "HFPO.DA" | Analyte == "PFOA" | Analyte == "PFOS") %>%
#   group_by(Site, Sample.Date)%>%
#   summarise(sum_pfas = sum(ppt))%>%
#   filter (sum_pfas > 130)
# 
# unique(Total_PFAS_Source_Over130$Site)
# 
#Plotting sites above 130
source_total_pfas_130_plot <-
  ggplot(Total_PFAS_WWTP_Over130, aes(x = Sample.Date, y = TotalPFAS, color = Site)) +
  geom_point(alpha = 0.7, size = 5)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  theme(axis.text.x = element_text(size = 7))+
  geom_hline(yintercept = 130, lty = 2)+
  scale_y_continuous(breaks=c(130, 200, 400, 600, 800, 1200))+
  theme(legend.position = "right",
        legend.text = element_text(size = 5), legend.title = element_text(size = 12))+
  xlab("Sample Date")+
  ylab("Total PFAS (ppt)")

print(source_total_pfas_130_plot + ggtitle("WWTP Sites with Total PFAS Above 130 ppt"))

```

## Question 2: Source Data ND/Qualifiers

### Question 2a: [Tay] Summary Stats: ND below 5; ND below 10; ND below 20 (distrubtion bell curve)

```{r}

# How many samples had reporting limits above health advisory levels? 
# maybe do >10 or by analyte

source_ND_under5 <-
  PFAS_Source_clean[!is.na(PFAS_Source_clean$lab.qual),] %>%
  filter(lab.qual <= 5)

source_ND_under10 <-
  PFAS_Source_clean[!is.na(PFAS_Source_clean$lab.qual),] %>%
  filter(lab.qual <= 10)

source_ND_under20 <-
  PFAS_Source_clean[!is.na(PFAS_Source_clean$lab.qual),] %>%
  filter(lab.qual <= 20)

summary(source_ND_under5)
summary(source_ND_under10)
summary(source_ND_under20)

ggplot(source_ND_under5)+
  geom_histogram(aes(x=lab.qual))+
  xlab("ND under 5 ppt")+
  ylab("# of Samples")

ggplot(source_ND_under10)+
  geom_histogram(aes(x=lab.qual))+
  xlab("ND under 10 ppt")+
  ylab("# of Samples")

ggplot(source_ND_under20)+
  geom_histogram(aes(x=lab.qual))+
  xlab("ND under 20 ppt")+
  ylab("# of Samples")


# histogram number of samples by all ND

ggplot(PFAS_Source_clean) +
  geom_histogram(aes(x=lab.qual))+
  xlim(0,600) +
  ylim(0,30) +
  xlab("ND ppt")+
  ylab("# of Samples")
```



### Question 2b: [Tay] Percentages (how many ND / total detections under 30) within any. Within GenX. See how many of which analytes have the worst percentage ratios --> play with those trends

```{r}

# of all the ND, how many are under 5

# how many total
source_ND <-
  PFAS_Source_clean[!is.na(PFAS_Source_clean$lab.qual),] %>%
  filter(lab.qual > 0)

# percentage under 5

source_nd_percent = (dim(source_ND_under5)[1] / dim(source_ND)[1])*100
source_nd_percent

# percentage under 20
source_nd_percent = (dim(source_ND_under20)[1] / dim(source_ND)[1])*100
source_nd_percent

```


## Question 3: WWTP Analytes

### Question 3a: PFOA, PFOS, GenX, sum PFAS accross sites

```{r}

# tried facet wrap

wwtp_pfoa.pfos.genx <-
  PFAS_WWTP_clean %>%
  filter(Analyte %in% c("PFOA", "PFOS", "HFPO.DA"))

# add month column
wwtp_pfoa.pfos.genx$Sample.Date <-
  ymd(wwtp_pfoa.pfos.genx$Sample.Date)

wwtp_pfoa.pfos.genx <- 
  mutate(wwtp_pfoa.pfos.genx, month = month(Sample.Date)) 

#facet wrap by month

ggplot(wwtp_pfoa.pfos.genx, aes(x=Analyte, y=ppt)) +
  geom_bar(position="stack", stat="identity") +
  facet_grid(Site ~ month)

# fewer sites


### TRY 2 -- need to convert HFPO-DA to HFPO.DA and run again

ggplot(wwtp_pfoa.pfos.genx, aes(x = Site, y = ppt))+
  geom_col(aes(fill = Analyte)) +
  labs(fill = "Analyte") +
  ylim(0,4500) +
  xlab("WWTP Site") +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

```


### Question 3b: Top analytes (aka are there analytes we aren't looking at)
```{r}

# plot analytes by site shows the only analytes worth knowing

ggplot(subset(PFAS_long_clean, Type == "WWTP"), aes(fill=Analyte, y=ppt, x=Site)) +     geom_bar(position="stack", stat="identity") +
    labs(fill = "Analyte") +
    theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  xlab("WWTP Sites")

# summarise  by column and get highest mean value

mean_analyte <-
  colMeans(PFAS_wide_clean[sapply(PFAS_wide_clean, is.numeric)], na.rm=TRUE) 

mean_analyte <- data.frame(mean_analyte)

mean_analyte <-
  mean_analyte %>%
  filter(lakename == "Paul Lake" | lakename == "Peter Lake")

```


### Question 3d: Long vs Short

```{r}


# group by chain length in Q1
# pfas_chain_group <-
  #PFAS_long_clean %>%
  #drop_na(ppt) %>%
  #group_by(Type, Site, chain.length) %>%
  #summarise(ppt = sum(ppt))

# bar plot stacked
ggplot(subset(pfas_chain_group, Type == "WWTP"), aes(x = Site, y = ppt))+
  geom_col(aes(fill = chain.length)) +
  labs(fill = "Chain Length") +
  xlab("WWTP Site") +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank())


```

### Question 3e: [Tay] Are any spots with one analyte over 70? over 30?

```{r}
wwtp_analyte_above_30 <- 
  PFAS_WWTP_clean %>%
  filter(ppt > 30)

wwtp_analyte_above_70 <- 
  PFAS_WWTP_clean %>%
  filter(ppt > 70)

source_analyte_above_30 <- 
  PFAS_Source_clean %>%
  filter(ppt > 30)

source_analyte_above_70 <- 
  PFAS_Source_clean %>%
  filter(ppt > 70)

unique(wwtp_analyte_above_30$Site) #27 Sites above 30 
unique(wwtp_analyte_above_70$Site) #13 sites above 70

unique(source_analyte_above_30$Site) #11 Sites above 30 
unique(source_analyte_above_70$Site) #5 sites above 70





#Plotting WWTP anayltes over the ppt standards
wwtp_analyte_30_plot <- 
  ggplot(wwtp_analyte_above_30, aes(x = Sample.Date, y = ppt, color = Site)) +
  geom_point(alpha = 0.5, size = 3)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  theme(axis.text.x = element_text(size = 7))+
  geom_hline(yintercept = 30, lty = 2)+
  scale_y_continuous(breaks=c(30, 100, 200, 400, 800, 1000))+
  theme(legend.position = "right", 
        legend.text = element_text(size = 5), legend.title = element_text(size = 12))+
  xlab("Sample Date")
print(wwtp_analyte_30_plot + ggtitle("WWTP sites with Analytes above 30 ppt"))

wwtp_analyte_70_plot <- 
  ggplot(wwtp_analyte_above_70, aes(x = Sample.Date, y = ppt, color = Site)) +
  geom_point(alpha = 0.5, size = 3)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  theme(axis.text.x = element_text(size = 7))+
  geom_hline(yintercept = 70, lty = 2)+
  scale_y_continuous(breaks=c(70, 200, 400, 800, 1000))+
  theme(legend.position = "right", 
        legend.text = element_text(size = 5), legend.title = element_text(size = 12))+
  xlab("Sample Date")
print(wwtp_analyte_70_plot + ggtitle("WWTP sites with Analytes above 70 ppt"))
```

### Question 3f: [Tay] Total PFAS over 130

```{r}


Total_PFAS_WWTP_Over130 <- 
  PFAS_WWTP_wide_clean %>%
  filter(TotalPFAS > 130)


# #WWTP by Site and Date with sum PFAS above 130
# Total_PFAS_WWTP_Over130 <- 
#   PFAS_long_clean[!is.na(PFAS_long_clean$ppt),] %>%
#   filter (Type == "WWTP") %>%
#   filter(Analyte == "PFBA"| Analyte == "PFBS" | Analyte == "PFDS" | Analyte == "PFDA" | Analyte == "PFDoA" | Analyte == "PFHpS" | Analyte == "PFHpA" | Analyte == "PFHxS" | Analyte == "PFHxA" | Analyte == "PFNS" | Analyte == "PFNA" | Analyte == "PFOSA" | Analyte == "N.EtFOSAA" | Analyte == "N.MeFOSAA" | Analyte == "PFPeS" | Analyte == "PFPeA" | Analyte == "PFTeA" | Analyte == "PFTriA" | Analyte == "PFUnA" | Analyte == "HFPO.DA" | Analyte == "PFOA" | Analyte == "PFOS") %>%
#   group_by(Site, Sample.Date)%>%
#   summarise(sum_pfas = sum(ppt))%>%
#   filter (sum_pfas > 130)
# 
# unique(Total_PFAS_Source_Over130$Site)
# 
#Plotting sites above 130
# wwtp_total_pfas_130_plot <-
#   ggplot(Total_PFAS_WWTP_Over130, aes(x = Sample.Date, y = sum_pfas, color = Site)) +
#   geom_point(alpha = 0.7, size = 5)+
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))+
#   theme(axis.text.x = element_text(size = 7))+
#   geom_hline(yintercept = 130, lty = 2)+
#   scale_y_continuous(breaks=c(130, 400, 600, 800, 1200, 1800, 2400))+
#   theme(legend.position = "right",
#         legend.text = element_text(size = 6), legend.title = element_text(size = 12))+
#   xlab("Sample Date")+
#   ylab("Total PFAS (ppt)")
# print(wwtp_total_pfas_130_plot + ggtitle("WWTP Sites with Total PFAS Above 130 ppt"))



#Plotting sites above 130
wwtp_total_pfas_130_plot <-
  ggplot(Total_PFAS_WWTP_Over130, aes(x = Sample.Date, y = TotalPFAS, color = Site)) +
  geom_point(alpha = 0.7, size = 5)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  theme(axis.text.x = element_text(size = 7))+
  geom_hline(yintercept = 130, lty = 2)+
  scale_y_continuous(breaks=c(130, 400, 600, 800, 1200, 1800, 2400))+
  theme(legend.position = "right",
        legend.text = element_text(size = 6), legend.title = element_text(size = 12))+
  xlab("Sample Date")+
  ylab("Total PFAS (ppt)")

print(wwtp_total_pfas_130_plot + ggtitle("WWTP Sites with Total PFAS Above 130 ppt"))
```





## Question 4: WWTP ND/Qualifiers

### Question 4a: [Tay] Summary Stats: ND below 5; ND below 10; ND below 20 (distrubtion bell curve)

```{r}
wwtp_ND_under5 <-
  PFAS_WWTP_clean[!is.na(PFAS_WWTP_clean$lab.qual),] %>%
  filter(lab.qual < 5)

wwtp_ND_under10 <-
  PFAS_WWTP_clean[!is.na(PFAS_WWTP_clean$lab.qual),] %>%
  filter(lab.qual < 10)

wwtp_ND_under20 <-
  PFAS_WWTP_clean[!is.na(PFAS_WWTP_clean$lab.qual),] %>%
  filter(lab.qual < 20)

summary(wwtp_ND_under5)
summary(wwtp_ND_under10)
summary(wwtp_ND_under20)

ggplot(wwtp_ND_under5)+
  geom_histogram(aes(x=lab.qual))+
  xlab("ND under 5 ppt")+
  ylab("# of Samples")

ggplot(wwtp_ND_under10)+
  geom_histogram(aes(x=lab.qual))+
  xlab("ND under 10 ppt")+
  ylab("# of Samples")

ggplot(wwtp_ND_under20)+
  geom_histogram(aes(x=lab.qual))+
  xlab("ND under 20 ppt")+
  ylab("# of Samples")


# histogram number of samples by all ND

ggplot(PFAS_WWTP_clean) +
  geom_histogram(aes(x=lab.qual))+
  xlim(0,600) +
  ylim(0,30) +
  xlab("ND ppt")+
  ylab("# of Samples")

```

### Question 4b: [Tay] Percentages (how many ND / total detections under 30) within any. Within GenX. See how many of which analytes have the worst percentage rations --> play with those trends

```{r}

# of all the ND, how many are under 5

# how many are under 5
dim(wwtp_ND_under5) #855

# how many total
wwtp_ND <-
  PFAS_WWTP_clean[!is.na(PFAS_WWTP_clean$lab.qual),] %>%
  filter(lab.qual > 0)

dim(wwtp_ND) #1259

# percentage

wwtp_nd_percent = (dim(wwtp_ND_under5)[1] / dim(wwtp_ND)[1])*100
wwtp_nd_percent

```

### Question 4c: [Tay] ND: Variability of lab qualifiers (know the level is above/below U) - by analyte per WWTP

```{r}

```



### LIT REVIEW 

# Total PFAS date v ppt by site (lines or facet)

```{r}

PFAS_WWTP_wide_clean$Sample.Date <- ymd(PFAS_WWTP_wide_clean$Sample.Date)

ggplot(PFAS_WWTP_wide_clean, aes(x=Sample.Date, y = TotalPFAS, group=Site)) +
  geom_line(aes(color=Site)) +
  geom_point(aes(color=Site)) +
  ylim(0,600) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %y")

# or by month column 

PFAS_WWTP_wide_clean <-
  PFAS_WWTP_wide_clean %>%
  mutate(month = month(Sample.Date))

ggplot(PFAS_WWTP_wide_clean, aes(x=month, y = TotalPFAS, group=Site)) +
  geom_line(aes(color=Site)) +
  geom_point(aes(color=Site)) +
  scale_x_continuous(breaks=c(7, 8, 9)) +
  ylab("Total PFAS (ppt)") +
  xlab("2019 Month") +
  theme(legend.position="right")

```

# Analytes by site (bar) with line at 70

```{r}

ggplot(PFAS_WWTP_clean, aes(fill=Analyte, x=Site, y=ppt)) +
  geom_bar(position="stack", stat="identity") +
  labs(fill = "Analyte") +
  geom_hline(yintercept = 70, lty = 2) +
  ylim(0,5000) +
  ylab("Total PFAS (ppt)") +
  xlab("WWTP Site") +
  geom_text(x = "City.of.Graham", y = 45, label = "70 ppt", hjust = 1, fontface = "bold")  +
  theme(axis.text.x = element_text(angle = 90,  hjust = 1),
    legend.position="right")


ggplot(subset(PFAS_Source_clean, !is.na(ppt)), aes(fill=Analyte, x=Site, y=ppt)) +
  geom_bar(position="stack", stat="identity") +
  labs(fill = "Analyte") +
  geom_hline(yintercept = 70, lty = 2) +
  ylim(0,3000) +
  ylab("Total PFAS (ppt)") +
  xlab("Source Site") +
  geom_text(x = "NORTHEASTCRKATSR1731OKELLYCHURCHRDNRDURHAM", y = 45, label = "70 ppt", hjust = 1, fontface = "bold")  +
  theme(axis.text.x = element_text(angle = 90,  hjust = 1),
    legend.position="right")

```

# Compare PFOA/PFOS and Total (stacked bar)

```{r}

PFAS_WWTP_wide_clean <-
  PFAS_WWTP_wide_clean %>%
  mutate(nonPFOA.PFOS = (TotalPFAS-SumPFOA.PFOS))

WWTP_sum.total <-
  PFAS_WWTP_clean %>%
  filter(Analyte %in% c("PFOA", "PFOS", "PFBA", "PFHxA", "PFNA", "PFBS", "PFHpA", "PFPeA"))

cols <- c(PFOA="blue", PFOS="blue", PFBA="black", PFHxA="black", PFNA="black", PFBS="black", PFHpA="black", PFPeA="black")

ggplot(WWTP_sum.total, aes(x=Site, y=ppt, fill=Analyte)) +
  geom_bar(stat="identity", position = "dodge") +
  scale_fill_manual(values = cols) +
  xlab("WWTP Site")+
  theme(axis.text.x = element_text(angle = 90,  hjust = 1),
    legend.position="right")
```

